Compositional Certified Resource Bounds

### 2015-04-13 paper/main.tex
remove + 0.|[y,x]| in figure 1
add a note on non-linear operations

\begin{figure}[t]
  \centering
\vspace{-1.5ex}
\begin{lstlisting}[mathescape]
${\color{blue}%
  \{ \cdot; \, 0 + \frac{T}{K} {\cdot} |[x,y]| + 0 {\cdot} |[y,x]| \} }$
while (x+K<=y) {
  ${\color{blue}%
  \{ x + K \le y; \; 0 + \frac{T}{K} {\cdot} |[x,y]| + 0 {\cdot} |[y,x]| \} }$
  x=x+K;
  ${\color{blue}%
  \{ x \leq y; \, T + \frac{T}{K} {\cdot} |[x,y]| + 0 {\cdot} |[y,x]| \} }$
  tick(T);
  ${\color{blue}%
  \{ x \leq y; \, 0 + \frac{T}{K} {\cdot} |[x,y]| + 0 {\cdot} |[y,x]| \} }$
}
${\color{blue}%
 \{ x \geq y; 0 + \frac{T}{K} {\cdot} |[x,y]| + 0 {\cdot} |[y,x]| \} }$
\end{lstlisting}
  \caption{Derivation of a tight bound on the number of
    ticks for a standard \emph{for loop}.  The parameters $K>0$ and
    $T>0$ are not program variables but denote concrete constants.
    % The reasoning works uniformly for all such $K$ and
    % $T$.
  }
  \label{fig:ex1}
\end{figure}

@@

In the potential functions, we treat constants as global variables
that cannot be assigned to.  For example, if the program contains the
constant $8128$ then we have a variable $c_{8128}$ and
$\state(c_{8128}) = 8128$.  We assume that every program state
includes the constant $c_0$.

@@

We describe the automatic amortized analysis for a subset of
expressions of Clight.  Clight is the first intermediate
language of the CompCert compiler~\cite{CompcertCACM}, it
is a cleaner C with a unified looping construct and
side-effect free expressions.
Assignments are restricted to the form $x \gets
y$ or $x \gets x \pm y$.  In the implementation, a Clight
program is converted into this form prior to analysis without changing
the resource cost.  This is achieved by using a series of \emph{cost-free
  assignments} that do not result in additional cost in the
semantics.  Non-linear operations such as $x \gets z*y$ or $x \gets
a[y]$ are handled by assigning $0$ to coefficients
like $q_{xa}$ and $q_{ax}$ that contain $x$ after the assignment.

>>>>

\begin{figure}[t]
  \centering
\vspace{-1.5ex}
\begin{lstlisting}[mathescape]
${\color{blue}%
  \{ \cdot; \, 0 + \frac{T}{K} {\cdot} |[x,y]| \} }$
while (x+K<=y) {
  ${\color{blue}%
  \{ x + K \le y; \; 0 + \frac{T}{K} {\cdot} |[x,y]| \} }$
  x=x+K;
  ${\color{blue}%
  \{ x \leq y; \, T + \frac{T}{K} {\cdot} |[x,y]| \} }$
  tick(T);
  ${\color{blue}%
  \{ x \leq y; \, 0 + \frac{T}{K} {\cdot} |[x,y]| \} }$
}
${\color{blue}%
 \{ x \geq y; 0 + \frac{T}{K} {\cdot} |[x,y]| \} }$
\end{lstlisting}
  \caption{Derivation of a tight bound on the number of
    ticks for a standard \emph{for loop}.  The parameters $K>0$ and
    $T>0$ are not program variables but denote concrete constants.
    % The reasoning works uniformly for all such $K$ and
    % $T$.
  }
  \label{fig:ex1}
\end{figure}

@@

In the potential functions, we treat constants as global variables
that cannot be assigned to.  For example, if the program contains the
constant $1988$ then we have a variable $c_{1988}$ and
$\state(c_{1988}) = 1988$.  We assume that every program state
includes the constant $c_0$.

@@

We describe the automatic amortized analysis for a subset of
expressions of Clight.  Clight is the first intermediate
language of the CompCert compiler~\cite{CompcertCACM}, it
is a cleaner C with a unified looping construct and
side-effect free expressions.
Assignments are restricted to the form $x \gets
y$ or $x \gets x \pm y$.  In the implementation, a Clight
program is converted into this form prior to analysis without changing
the resource cost.  This is achieved by using a series of \emph{cost-free
  assignments} that do not result in additional cost in the
semantics.  Non-linear operations such as $x \gets z*y$ or $x \gets
a[y]$ are handled by assigning $0$ to coefficients
like $q_{xa}$ and $q_{ax}$ that contain $x$ after the assignment.
This sound treatment ensures that no further loop bounds
depend on the result of the non-linear operation.
